---
title: "R Notebook"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

## **Instructions**

1. Write your code to each question in R script file. The answers should be correctly
ordered.
2. When the question is asking more than just writing an R code, write your answer to
each question in Word document. In addition, if the question is asking you for the
visualization, copy and paste your plots (or figures) to the Word document. The answer
should be correctly ordered.
3. After you finish all the problems, upload both your R script file and Word document to
Homework 1 Dropbox in the course Elearning.

Use set.seed(210) at the beginning of Problem 1 and Problem 2.

**Problem 1**. The Pareto distribution has been used in economics as a model for a density function
with a slowly decaying tail. The Pareto(a,b) distribution has pdf:

$$ f(x; a, b) = \frac{a b^a}{x^{a + 1}}, \; b > 0, \; a > 0, \; x \ge b $$

1. Use the inverse transform method to generate a random sample of size 1000 from the
Pareto(3,2) distribution.
2. Plot the density histogram of the sample.
3. Add the theoretical density curve f (x) to the density histogram.

First, we find the CDF from the pdf given.
F(x) = P(X <= x) = $$F(x) = \int_0^x\frac{ab^a}{t^{a+1}}dt=-b^ax^{-a}$$

For the Pareto(3, 2) distribution we have
$$ f(x; 3, 2) = \frac{3 * 2^3}{x^{3+1}} = \frac{24}{x^{4}}, x \gt 2$$
$$F(x) =  \frac{-8}{x^3}$$

Next, we find the inverse
$$F^{-1}(u) = \frac{-2u^{2/3}}{u}$$

```{r}
set.seed(210)
n <- 1000
u <- runif(n)
x <- (-2*u^(2/3))/(u)
hist(x, prob = T, xlab = "x", main = "Pareto(3,2)", freq = F)
y <- seq(0, 1, 0.01)
lines(y, 24/y^4)
#curve(24/(x^4), col = "red", add = T)
```

4. Write your own rpareto function that simulates random numbers having a Pareto distribution with specified parameters. (Hint: rpareto function requires three arguments;sample size, shape parameter, and scale parameter.)
```{r}
rpareto <- function(size, shape, scale){
  set.seed(210)
  x <- vector(length=size)
  u <- runif(size)
  x <- (-scale*u^(scale/shape))/(u)
  return(x)
}
```

5. Use your rpareto function to generate a random sample of size 1000 from the Pareto(2,1)
distribution and find a sample mean.
```{r}
sample <- rpareto(1000, 2, 1)
sum(sample)/length(sample)
mean(sample)
```




