---
title: "R Notebook"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

## **Instructions**

1. Write your code to each question in R script file. The answers should be correctly
ordered.
2. When the question is asking more than just writing an R code, write your answer to
each question in Word document. In addition, if the question is asking you for the
visualization, copy and paste your plots (or figures) to the Word document. The answer
should be correctly ordered.
3. After you finish all the problems, upload both your R script file and Word document to
Homework 1 Dropbox in the course Elearning.

Use set.seed(210) at the beginning of Problem 1 and Problem 2.

**Problem 1**. A discrete random variable X has probability mass function (pmf)

| x      | 0     | 1     | 2     | 3     |
| ------ | ------|------ | ------|------ |  
| P(X=x) | 0.064 | 0.288 | 0.432 | 0.216 |

1. Use the inverse transform method to generate a random sample of size 1000 from the distribution of X.

```{r}
set.seed(210)
n <- 1000
x <- rep(0, n)
u <- runif(n)
F_x = c(0.064, 0.288, 0.432, 0.216)
for (i in 1:n){
  if (u[i] <= F_x[1]){
    x[i] <- 0
  }
  else if (F_x[1] < u[i] & u[i] <= (F_x[1] + F_x[2])){
    x[i] <- 1
  }
  else if ((F_x[1] + F_x[2]) < u[i] & u[i] <= (F_x[1] + F_x[2] + F_x[3])){
    x[i] <- 2
  }
  else if ((F_x[1] + F_x[2] + F_x[3]) < u[i] & u[i] <= 1){
    x[i] <- 3
  }
}
table(x)
mean(x)
```

2. Calculate relative frequency for each x and compare with the theoretical probabilities given in the pmf table. For example, to find the relative frequency for 0, count how many times 0 appeared and divide by the sample size 1000

```{r}
table(x)/length(x)
```
