---
title: "Homework 4"
output:
  word_document: default
  html_notebook: default
---

## **Instructions**

1. Write your code to each question in R script file. The answers should be correctly
ordered.
2. When the question is asking more than just writing an R code, write your answer to
each question in Word document. In addition, if the question is asking you for the
visualization, copy and paste your plots (or figures) to the Word document. The answer
should be correctly ordered.
3. After you finish all the problems, upload both your R script file and Word document to
Homework 1 Dropbox in the course Elearning.


```{r}
set.seed(318)
```

**Problem 1** Let X = the number of heads obtained from flipping a fair coin 4 times. Here, we have that there 4 trials, "success" is defined as getting a head face-up with probability 0.5 for each trial, and the outcome of one trial doesn't affect another.

1 . Find the Probability Mass Function (pmf) table for random variable X. (Hint: X follows a binomial random variable. You may use dbinom() function to find corresponding probability for each outcome).

```{r}
S <- seq(0, 4)
pmf <- c (dbinom(S, 4, 0.5))
table(pmf, S)
```

2. Use the Metropolis algorithm to simulate X, 50000 times.
```{r}
n <- 50000
x <- rep(0, n)
Q <- matrix(c(.25, .1875, .1875, .1875, .1875, .1875, .25, .1875,.1875,.1875, .1875, .1875, .25, .1875,.1875,.1875,.1875,.1875,.25, .1875, .1875, .1875, .1875, .1875,.25), nrow=5, ncol=5, byrow = TRUE)
x[1] <- 2
for(i in 1:(n-1)){
	# pick new sample
	xnew <- sample(S, 1, prob = Q[which(S == x[i]), ])
	# acceptance condition
	u <- runif(1)
	if (u < pmf[which(S == xnew)]/pmf[which(S == x[i])]){
		x[i+1] <- xnew
	}
	else{
		x[i+1] <- x[i]
	}
}

table(x)
table(x)/n
```

3. Find the Monte Carlo estimate of E(X)

```{r}
mean(x)
```

4. Find the Monte Carlo estimate of Var(X)
```{r}
sd(x)^2
var_x = mean(x^2)-mean(x)^2 
var_x
```

**Problem 2** The Cauchy distribution with location l and scale s has density function
f(x) = 1 / (π s (1 + ((x-l)/s)^2))

1. Use the Metropolis algorithm to generate a random sample size of 10000 from the Cauchy distribution (s=2, l=0). Hint: Use normal distribution with a mean value at the current sample and a standard deviation of 10 as a proposal distribution
```{r}
set.seed(318)
n2 <- 10000
x2 <- rep(0, n2)
f_x <- function(x){
  dcauchy(x, location = 0, scale = 2)
}
q <- function(x){
	rnorm(1, x, 10)
}
x2[1] = 2
for(i in 1:(n2-1)){
	xnew <- q(x2[i])
	u <- runif(1)
	if (u < f_x(xnew)/f_x(x2[i])){
		x2[i+1] <- xnew
	}
	else{
		x2[i+1] <- x2[i]
	}
}

hist(x2, prob = TRUE, ylim=c(0,.15))
y <- seq(-20, 20, 1)
lines(y, f_x(y), col = "red")

mean(x)
```

2. Compute the sample percentiles and compare with the Cauchy distribution(scale = 2, location = 0) percentiles. Does the sample we generated follow the Cauchy distribution?

```{r}
p <- c(.10, .25, .5, .75, .90)
percentiles <- qcauchy(p, location = 0 ,scale=2, log.p = FALSE)
table(percentiles, p)
```

