---
title: "R Notebook"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

## **Instructions**

1. Write your code to each question in R script file. The answers should be correctly
ordered.
2. When the question is asking more than just writing an R code, write your answer to
each question in Word document. In addition, if the question is asking you for the
visualization, copy and paste your plots (or figures) to the Word document. The answer
should be correctly ordered.
3. After you finish all the problems, upload both your R script file and Word document to
Homework 1 Dropbox in the course Elearning.

Use set.seed(210) at the beginning of Problem 1 and Problem 2.

**Problem 1**. 
The normal distribution is one of the most important distribution in statistics
because the nature provides numerous examples of populations of measurements that, at
least approximately, follow a normal distribution.

1. Use the rejection sampling method to generate a random sample of size 1000 from the
standard normal distribution (mean = 0 and sd = 1). Use a standard Cauchy distribution (location = 0 and scale = 1) as a proposal distribution g(y). You can generate
a random sample from the standard Cauchy distribution using rcauchy(). Moreover,
you can calculate normal density, f(y) and Cauchy density, g(y) using dnorm() and
dcauchy() respectively. Choose c such that f(y) ≤ cg(y) and report your choice of c

2. Compute the sample percentiles from the sample we obtained in previous problem and
compare with the standard normal distribution percentiles. You can obtain standard
normal distribution percentiles using qnorm().

```{r}
set.seed(210)
n <- 1000
f_x <- rnorm(n, 0, 1)
g_x <- rcauchy(n, 0, 1)
M <- max(f_x/g_x)
range <- seq(1, 1000, 1)
plot(range, f_x, type = 'l', ylim=c(-5, 5), col='red')
lines(range, g_x, type='l', col='blue')
k <- 0 #counter for accepted
j <- 0 #counter for iterations
x <- rep(0, n)
while(k < n){
  y <- rcauchy(1)
  u <- runif(1)
  j <- j + 1
  if(u<dnorm(y)/(dcauchy(y)*M)){
    k <- k + 1
    x[k] <- y
  }
}
p <- seq(0.1, 0.9, 0.01)
qhat <- quantile(x, p)
Q <- qnorm(p, 0, 1)
round(rbind(qhat, Q), 3)
plot(Q, qhat)

```
**Problem 2**.
1. Use the importance sampling method to estimate Ef (X). (Hint: you may use normal
distribution as your importance function and use rnorm() function to draw samples.)
2. Draw a plot to show the convergence of the importance sampling approximation.

```{r}
n <- 1000

#x<- rnorm(n, mean=0, sd=1)

g <- function(x){
  (1/2*(1/sqrt(2*pi)*exp(-((x-1)^2)/2))) + (1/2*(1/sqrt(2*pi)*exp(-((x-3)^2)/2)))
}

x1 <- rnorm(n, mean = 1, sd = 1)
fg1 <-x1*g(x1)/dnorm(x1, mean = 1, sd = 1)
theta.hat1 <- mean(fg1)
theta.hat1


x2 <- rnorm(n, mean = 3, sd = 1)
fg2 <- x2*g(x2)/dnorm(x2, mean = 3, sd = 1)
theta.hat2 <- mean(fg2)
theta.hat2

x3 <- rnorm(n, mean = 0, sd = 1)
fg3 <- x3*g(x3)/dnorm(x3, mean = 0, sd = 1)
theta.hat3 <- mean(fg3)
theta.hat3
par(mfrow =c(1,3))

plot(cumsum(fg1)/1:n, type = "l", main = 'X ~ N(1, 1)')
abline(a=theta.hat1, b=0, col='red')

plot(cumsum(fg2)/1:n, type = "l",main = 'X ~ N(3, 1)')
abline(a=theta.hat2, b=0, col='red')

plot(cumsum(fg3)/1:n, type = "l", main = 'X ~ N(0, 1)')
abline(a=theta.hat3, b=0, col='red')
```



